---
title: "creating_new_communication_survey_variables. Any opinions and conclusions expressed herein are those of the author(s) and do not reflect the views of the U.S. Census Bureau. The Census Bureau has reviewed this data product to ensure appropriate access, use, and disclosure avoidance protection of the confidential source data (Project No. P-7529180, Disclosure Review Board (DRB) approval number:  CBDRB-FY24-EWD001-007)"
output: html_document
date: "2024-08-13"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#loading packages
```{r}
library(tidyverse)
library(haven)
library(lubridate)
library(ggplot2)
library(DBI)
library(stringr)
library(vtable)
```

# Getting Burn Rate and other variables for the compendium (new variables)

file paths have been changed for external release and no longer work
```{r}
# read in tables
comm = read_csv("inputs/communication.csv")
survey = read_csv("inputs/survey.csv")
call = read_csv("inputs/call.csv")
email = read_csv("inputs/email.csv")
mail = read_csv("inputs/mail.csv")
robo = read_csv("inputs/robocall.csv")
```

```{r}
# merging communication and survey tables 
comm_survey = comm_survey %>%
  group_by(survey_ID) %>%
  mutate(days_past_last_comm = as.numeric(date - lag(date), unit = 'days'))

# creating new variables 
# days since open date on survey 
comm_survey$days_past_open <- as.numeric(difftime(comm_survey$date, comm_survey$open_date, units = "days"))

# days past last contact point  
comm_survey = comm_survey %>%
  group_by(survey_ID) %>%
  mutate(days_past_last_comm = as.numeric(date - lag(date), unit = 'days'))

# days past same type of communication
comm_survey %>%
  group_by(survey_ID, com_type) %>% 
  mutate(days_past_same_comm = as.numeric(date - lag(date), unit = 'days')) %>% 
  select(date, days_past_same_comm)

# reducing comm_survey back into communication table 
comm_more = comm_survey[c(1:9, 25:27)] 

# relabeling columns
names(comm_more)[names(comm_more) == 'feb_birth.x'] <- 'feb_birth'
names(comm_more)[names(comm_more) == 'may_birth.x'] <- 'may_birth'
names(comm_more)[names(comm_more) == 'august_birth.x'] <- 'august_birth'
comm_more


# collection period length to survey table
survey$collection_length = as.numeric(survey$close_date - survey$open_date, unit = 'days')


```

Adding day of the week for email table
```{r}
# adding day of week 
email$day_of_week[email$day == 1] = "Sunday"
email$day_of_week[email$day == 2] = "Monday"
email$day_of_week[email$day == 3] = "Tuesday"
email$day_of_week[email$day == 4] = "Wednesday"
email$day_of_week[email$day == 5] = "Thursday"
email$day_of_week[email$day == 6] = "Friday"
email$day_of_week[email$day == 7] = "Saturday"
```


Loading records of individual burn records 

database connection no longer connects/connection has been changed for external publication
```{r}
#
# connect to database 
pass = rstudioapi::askForPassword("Database password")
con_string = paste0(
  "Driver=driver;DBQ=db;SVD=EXTRACTS;UID=username;PWD=",pass)
# aces
query_aces <- 'select * from ACES'
con = dbConnect(odbc::odbc(), .connection_string = con_string, timeout = 10)
aces_query = dbplyr::build_sql(dbplyr::sql(query_aces), con = con)
aces = DBI::dbGetQuery(con, aces_query)
# getting br
aces_br = aces %>% group_by(statp) %>% 
  summarise(
            used = n_distinct(ID[auth_used == "U"& 
                                (!is.na(authdate_audit) & !is.na(authdate_bank))]), 
            total = n_distinct(ID), br = (used/total)*100 )
aces_br$year = as.numeric(substr(aces_br$statp, 1, 4))
aces_br$statp = "ACES"

# arts
query_arts <- 'select * from ARTS'
con = dbConnect(odbc::odbc(), .connection_string = con_string, timeout = 10)
arts_query = dbplyr::build_sql(dbplyr::sql(query_arts), con = con)
arts = DBI::dbGetQuery(con, arts_query)
# getting br
arts_br = arts %>% group_by(statp) %>% 
  summarise(
            used = n_distinct(ID[auth_used == "U" & 
                                (!is.na(authdate_audit) & !is.na(authdate_bank))]), 
            total = n_distinct(ID), br = (used/total)*100 )
arts_br$year = as.numeric(substr(arts_br$statp, 1, 4))
arts_br$statp = "ARTS"

# awts
query_awts <- 'select * from AWTS'
con = dbConnect(odbc::odbc(), .connection_string = con_string, timeout = 10)
awts_query = dbplyr::build_sql(dbplyr::sql(query_awts), con = con)
awts = DBI::dbGetQuery(con, awts_query)
# getting br
awts_br = awts %>% group_by(statp) %>% 
  summarise(
            used = n_distinct(ID[auth_used == "U" & 
                                (!is.na(authdate_audit) & !is.na(authdate_bank))]), 
            total = n_distinct(ID), br = (used/total)*100 )
awts_br$year = as.numeric(substr(awts_br$statp, 1, 4))
awts_br$statp = "AWTS"

# sas
query_sas <- 'select * from SAS'
con = dbConnect(odbc::odbc(), .connection_string = con_string, timeout = 10)
sas_query = dbplyr::build_sql(dbplyr::sql(query_sas), con = con)
sas = DBI::dbGetQuery(con, sas_query)
# getting br
sas_br = sas %>% group_by(statp) %>% 
  summarise(
            used = n_distinct(ID[auth_used == "U" & 
                                (!is.na(authdate_audit) & !is.na(authdate_bank))]), 
            total = n_distinct(ID), br = (used/total)*100 )
sas_br$year = as.numeric(substr(sas_br$statp, 1, 4))
sas_br$statp = "SAS"
sas_br

# cos/asm (kept in different data source)
c17 = read.delim("Bank Authcode/COSASM_2017.dat", sep = "|")
c18 = read.delim("Bank Authcode/COSASM_2018.dat", sep = "|")
c19 = read.delim("Bank Authcode/COSASM_2019.dat", sep = "|")
c20 = read.delim("Bank Authcode/COSASM_2020.dat", sep = "|")
c21 = read.delim("Bank Authcode/COSASM_2021.dat", sep = "|")
c22 = read.delim("Bank Authcode/COSASM_2022.dat", sep = "|")
c_list =  lapply(c("c17", 'c18', 'c19', 'c20', 'c21', 'c22'), get)
cosasm = do.call(rbind, c_list)
names(cosasm)[names(cosasm) == "DATE"] = "AUTHDATE_BANK"
cosasm$AUTH_USED = substr(cosasm$STATUS, 1, 1)
cosasm$AUTHDATE_BANK <- as.Date(cosasm$AUTHDATE_BANK, format = "%d-%b-%Y")
# getting br
cosasm_br = cosasm %>% group_by(YEAR) %>% 
  summarise(
            used = n_distinct(COMPANYID[STATUS == "USED"]), 
            total = n_distinct(COMPANYID), br = (used/total)*100 )
cosasm_br$year = as.numeric(substr(cosasm_br$YEAR, 1, 4))
cosasm_br$statp = "COS/ASM"
cosasm_br = cosasm_br %>% select(c(statp, used, total, br, year))
cosasm_br

# m3ufo (kept in different data source)
m17 = read.delim("Bank Authcode/M3UO14_2017.dat", sep = "|")
m18 = read.delim("Bank Authcode/M3UO14_2018.dat", sep = "|")
m19 = read.delim("Bank Authcode/M3UO14_2019.dat", sep = "|")
m20 = read.delim("Bank Authcode/M3UO14_2020.dat", sep = "|")
m21 = read.delim("Bank Authcode/M3UO14_2021.dat", sep = "|")
m22 = read.delim("Bank Authcode/M3UO14_2022.dat", sep = "|")
m_list =  lapply(c("m17", 'm18', 'm19', 'm20', 'm21', 'm22'), get)
m3uo14 = do.call(rbind, m_list)
m3uo14
# calculating br
m3_br =  m3uo14 %>% group_by(YEAR) %>% 
  summarise(
            used = n_distinct(COMPANYID[STATUS == "USED"]), 
            total = n_distinct(COMPANYID), br = (used/total)*100 )
m3_br$year = as.numeric(substr(m3_br$YEAR, 1, 4))
m3_br$statp = "M3UFO"
m3_br = m3_br %>% select(c(statp, used, total, br, year))
m3_br

# combining all the burn rates and adding them to survey table 
br_list =  lapply(c("aces_br", 'awts_br', 'arts_br', 'm3_br', 'cosasm_br', 'sas_br'), get)
br_data = do.call(rbind, br_list)
br_data$ID =  paste(br_data$statp, br_data$year, sep = "_")
br_data = br_data %>% select(-c(statp, year))
br_data
survey_br = merge(survey, br_data, by = "ID") 

```

# Code for altering com_more

```{r}
comm_survey = comm_survey %>%
  group_by(survey_ID) %>%
  mutate(days_past_last_comm = as.numeric(date - lag(date), unit = 'days'))

comm_more = comm_survey %>% arrange(survey_ID, date) %>% group_by(survey_ID) %>%
  mutate(days_past_last_comm = as.numeric(date - lag(date), unit = 'days'))

comm_more = comm_more %>%
  arrange(survey_ID, date) %>% group_by(survey_ID,com_type) %>%
  mutate(days_past_same_comm = as.numeric(date - lag(date), unit = 'days')) 
```

```{r}
# writing new csv for added variables to the tables (not all tables had extra variables added)
write.csv(comm_more,"inputs/compendium updated/comm_more.csv", 
          row.names = FALSE)
write.csv(survey_br, "inputs/compendium updated/survey_more.csv",
          row.names = FALSE)
write.csv(call, 
          "inputs/compendium updated/call_more.csv",
          row.names = FALSE)
write.csv(email, "inputs/compendium updated/email_more.csv", 
          row.names = FALSE)
write.csv(mail, "inputs/compendium updated/mail_more.csv", 
          row.names = FALSE)
write.csv(robo, "inputs/compendium updated/robo_more.csv", 
          row.names = FALSE)
```

